---
title: "CCN Notes"
subtitle: "Numbering System for CCNs"
date: "`r Sys.time()`"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ccn)
```

```{r pkg-version}
utils::packageVersion("ccn")
```


> SEC 2779A, PG 403 (som107c02.pdf)

_Rev. 152, ISS: `03-25-16`, EFF: `04-04-16`, IMP: `04-04-16`_

__CCN__ replaced the terms _Medicare Provider Number_, _Medicare Identification Number_ 
and _OSCAR Number_ in 2007. It is used to verify Medicare/Medicaid 
certification for survey and certification, assessment-related activities and 
communications. Additionally, CMS data systems use the CCN to identify each 
individual provider or supplier that has or currently does participate in Medicare 
and/or Medicaid. The RO, not the SA or MAC, assigns the CCN and maintains adequate controls.

## Medicare Providers

CCNs for Providers and Suppliers paid under _Medicare Part A_ have __6 digits__.

The first 2 digits identify the State in which the provider is located. The last
4 digits identify the type of facility.

The following is a list of all State Codes:

```{r states}
ccn:::STATE$CODE |> 
   ccn:::make_df(c("State", "Code")) |> 
   head(10) |> 
   knitr::kable()
```

The last 4 digits identify the facility type, sequentially from
within the following range:

```{r care-range}
ccn:::MEDICARE$ABBR |> 
   ccn:::make_df(c("Facility", "Range")) |> 
   head(10) |> 
   knitr::kable()
```

_Examples_: __CMHCs in Florida__

```{r florida}
ccn("681400")
ccn("684702")
ccn("691400")
```

### Organ Procurement Organizations (OPOs)

OPOs are assigned a 6-digit alphanumeric CCN. The first 2 digits identify the
State Code. The third digit is the alpha character "P." The remaining 3 digits
are the unique facility identifier.

```{r opo}
ccn("01P001")
ccn("05P001")
```

## Medicare Suppliers

> Section 2779A2, Page 407

_Rev. 198, ISS: `01-17-20`, EFF: `01-17-20`, IMP: `01-17-20`_

Suppliers that are paid by Part B carriers have a 10-digit alphanumeric CCN. The
first 2 digits identify the State in which the supplier is located. The third
digit is an alpha character that identifies the type of facility. The remaining
7 digits are the unique facility identifier.

The following characters in the third position indicate:

   * C - Ambulatory Surgical Centers
   * D - Clinical Laboratory Improvement Amendments of 1988 (CLIA) Laboratories
   * X - Portable X-Ray Facilities

The last 7 digits of the CCN for the above suppliers will be within the number series
`0000001-9999999`.

_Examples_:

```{r supplier}
# ASC
ccn("10C0001062")
# CLIA
ccn("45D0634589")
# Portable X-Ray
ccn("21X0009807")
```

## Medicaid-Only Providers

> Section 2779B, Page 408

_Rev. 123, ISS: `10-03-14`, EFF: `10-03-14`, IMP: `10-03-14`_

For certification purposes, Title XIX-Only Providers are identified by a 6-digit
alphanumeric CCN. The first 2 digits identify the State in which the provider is
located. The third position, which is an alpha character, identifies the type of
facility by level or type of care being provided. The last 3 digits make up a
sequential number series beginning with 001.

```{r}
ccn("01L008")
```


```{r caid-range}
ccn:::MEDICAID$ABBR |> 
   ccn:::make_df(c("Facility", "Range")) |> 
   head(10) |> 
   knitr::kable()
```


> __EXCEPTION:__ As of October 1, 2019, an IPPS-excluded hospital is no longer
precluded from having an IPPS-excluded psychiatric and/or rehabilitation unit.
See section 2779C and 2779C1 for additional CCN numbering detail. 

> __Note:__ An IPPS-excluded hospital may not have an IPPS-excluded unit of the
same type (psychiatric or rehabilitation) as the hospital (for example, an
Inpatient Rehabilitation Facility (IRF) may not have an IRF unit).


# Sources

  - [CMS Manual](https://www.cms.gov/regulations-and-guidance/guidance/manuals/downloads/som107c02.pdf)
  - [ResDAC 1](https://resdac.org/cms-data/variables/full-cms-certification-number-provider)
  - [ResDAC 2](https://resdac.org/cms-data/variables/provider-number)
  - [The Joint Commission](https://manual.jointcommission.org/releases/TJC2025A/DataElem0276.html)
  - [QCOR](https://qcor.cms.gov/main.jsp)
  - [HHA Branch Identification Numbers](https://www.cms.gov/Regulations-and-Guidance/Guidance/Transmittals/Downloads/R2SOM.pdf)
